\section{OGRE3D}
\label{sec:OGRE3D}
OGRE is a scene-oriented, flexible 3D rendering engine (as opposed to a game 
engine) written in C++ designed to make it easier and intuitive for developers 
to produce applications utilising hardware-accelerated 3D graphics. The class 
library abstracts the details of using the underlying system libraries like 
Direct3D and OpenGL and provides an interface based on world objects and other 
high level classes.

The engine is free software, licensed under the LGPL and has a very active 
community. It has been used in some commercial games like Ankh and Pacific 
Storm. It was Sourceforge's project of the month in March 2005.

1.0.0 "Azathoth" was released in February 2005, the current release in the 
1.x.y series is 1.4.3 "Eihort" (July 2007).

\subsection{General}
\label{sec:OGREGeneral}
As its name states, OGRE is "just" a rendering engine. As thus, its main 
purpose is to provide a general solution for graphics rendering. Though it also 
comes with some facilities\footnote{vector and matrix classes, memory handling, et 
cetera} they must be thought as helpers only. It is not an all-in-one solution 
in terms of game development or simulation as it doesn't provide audio or 
physics support, for instance.

Generally, this is thought of as the main drawback of OGRE, but it could also 
be seen as a feature of the engine. The choice of OGRE as a graphics engine 
allows developers the freedom to use whatever physics, input, audio and other 
libraries they want and allows the OGRE development team to focus on graphics 
rather than distribute their efforts amongst several systems. OGRE explicitly 
supports the OIS, SDL and CEGUI libraries, and includes the Cg toolkit.

Currently OGRE is published under a dual license\footnote{One being LGPL, the other one 
called OGRE Unrestricted License (OUL)}, to make it possible to be chosen for 
console development as well, because most of the publishers reject using 
free/open-source software in that particular market.

\subsection{Features}
\label{sec:Features}
\begin{itemize}
	\item object oriented design
	\item plug-in architecture
	\item scene graph based
	\item support of a wide variety of scene managers
	\item fully multi-platform
	\item support of GLSL, HLSL, Cg and asm.
	\item full support for hardware weighted multiple bone skinning
	\item support for effects such as HDR, blooming, saturation, brightness, blurring and noise
	\item particle system with extensible rendering and customizable effectors and emitters
\end{itemize}

A full overview of the features provided by OGRE can be found on the web site http://www.ogre3d.org.